Discrete Probability for Discrete Quantum Computing

### 2016-05-04 prob.tex
prob.tex update the old [Dempster-Shafer Theory of Evidence] example...

\begin{example}[Dempster-Shafer Theory of Evidence] We have five
  employees whose precise age is not known. All is given is a range for
  each employee:
\[\begin{array}{rcl}
\textrm{Age}(M_{1}) & \in & [22,26]=D_{1}\\
\textrm{Age}(M_{2}) & \in & [20,22]=D_{2}\\
\textrm{Age}(M_{3}) & \in & [30,35]=D_{3}\\
\textrm{Age}(M_{4}) & \in & [20,22]=D_{4}\\
\textrm{Age}(M_{5}) & \in & [28,30]=D_{5}
\end{array}\]
What is the probability that an employee's age is in the range
$[20,25]$? Looking at the data, it is \emph{possible} for $M_1$'s age
to be within the range, it is \emph{not possible} for $M_3$'s age to
be within the range, and it is \emph{certain} or \emph{necessary} that
$M_4$'s age is in the range. Clearly saying that an event is
\emph{necessary} is equivalent to saying that its complement is
\emph{not possible}. Aggregating the results for the five employees,
we can calculate that is necessary that 2 employees have their ages
within the range, and it is possible that 3 employees have their ages
within the range. We express this formally as saying that the
probability of the event is $[0.4,0.6]$. Now consider another event
asking whether the ages are some other disjoint range
$[26,40]$. Reasoning in a similar way we calculate that the
probability for this event is $[0.4,0.6]$. Now let's take the two events
together and ask about the possibility of the ages to be in the range
$[20,40]$. Clearly that probability must be $[1,1]$ as every
employee's age is in that range. The way to calculate this probability
from the individual probabilities of the events is as follows: 
\begin{itemize}
\item \amr{can you try to fix this... perhaps it is not consistent
    with the interval-valued formalization but it is really nice if we
    can make it work}
\item Because ages are all integers, and we only consider ages from 20 to
40. Hence, the sample space~$\Omega$ can be $\left\{ 20,21,\ldots,39,40\right\} $,
and the set of events are $2^{\Omega}$ as usual.\\
\yutsung{I have numerically verified if we set $\Omega=\left\{ 20,22,25,26,28,30,35,40\right\} $,
and the answer should be the same as $\left\{ 20,21,\ldots,39,40\right\} $
because only end-points should matter in the question... I will numerically
verify $\left\{ 20,21,\ldots,39,40\right\} $, or prove they are the
same later...} 
\item Given a event~$E\subseteq\Omega$, the probability of $E$ can be
defined by 
\[
\pmeas(E)=\left[\frac{1}{5}\left|\left\{ i\middle|D_{i}\cap E\ne\emptyset\right\} \right|,\frac{1}{5}\left|\left\{ i\middle|D_{i}\subseteq E\right\} \right|\right]\textrm{ ,}
\]
where $\left|A\right|$ is the number of elements in $A$. Then, we
can verify that $\pmeas$ in an interval-value probability measure.
\end{itemize}
\qed\end{example}

>>>>

\begin{example}[Dempster-Shafer Theory of Evidence] We have five
  employees whose precise age is not known. All is given is a range for
  each employee:
\[\begin{array}{rcl}
\textrm{Age}(M_{1}) & \in & \{ 23,24 \}=D_{1}\\
\textrm{Age}(M_{2}) & \in & \{ 20,21,22\}=D_{2}\\
\textrm{Age}(M_{3}) & \in & \{ 20,21 \}=D_{3}
\end{array}\]
What is the probability that an employee's age is in the range
$\{22,23,24\}$? Looking at the data, it is \emph{possible} for $M_2$'s age
to be within the range, it is \emph{not possible} for $M_3$'s age to
be within the range, and it is \emph{certain} or \emph{necessary} that
$M_1$'s age is in the range. Clearly saying that an event is
\emph{necessary} is equivalent to saying that its complement is
\emph{not possible}. Aggregating the results for the three employees,
we can calculate that is necessary that 1 employees have their ages
within the range, and it is possible that 2 employees have their ages
within the range. We express this formally as saying that the
probability of the event is $\left[\frac{1}{3},\frac{2}{3}\right]$. Now consider
another event asking whether the ages are some other disjoint range
$\{20,21\}$. Reasoning in a similar way we calculate that the
probability for this event is $\left[\frac{1}{3},\frac{2}{3}\right]$. Now let's
take the two events together and ask about the possibility of the ages to be in the range
$\{20,21,22,23,24\}$. Clearly that probability must be $[1,1]$ as every
employee's age is in that range. This problem
looks puzzle if we want to attack it directly, but will be more clear
if we think in terms of conditional probability. We will compute the
conditional probability for the usual real-valued probability first
and for the interval-valued probability later.
\begin{itemize}
\item We randomly draw an employee among the three, so each employee has
probability $\frac{1}{3}$ to be drawn, i.e.,
\[
\pmeas\left(\left\{ i=1\right\} \right)=\pmeas\left(\left\{ i=2\right\} \right)=\pmeas\left(\left\{ i=3\right\} \right)=\frac{1}{3}
\]
In order to compute the usual real-valued probability, we need a probability
distribution within the possible range $D_{1}$, $D_{2}$, and $D_{3}$.
Let's assume they are equally probable, i.e., 
\[
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas\left(\left\{ \textrm{Age}(M_{1})=23\right\} \right) & = & \frac{1}{2}\\
\pmeas\left(\left\{ \textrm{Age}(M_{1})=24\right\} \right) & = & \frac{1}{2}
\end{array} & \begin{array}{rcl}
\pmeas\left(\left\{ \textrm{Age}(M_{2})=20\right\} \right) & = & \frac{1}{3}\\
\pmeas\left(\left\{ \textrm{Age}(M_{2})=21\right\} \right) & = & \frac{1}{3}\\
\pmeas\left(\left\{ \textrm{Age}(M_{2})=22\right\} \right) & = & \frac{1}{3}
\end{array}\end{array}\begin{array}{rcl}
\pmeas\left(\left\{ \textrm{Age}(M_{3})=20\right\} \right) & = & \frac{1}{2}\\
\pmeas\left(\left\{ \textrm{Age}(M_{3})=21\right\} \right) & = & \frac{1}{2}
\end{array}
\]
Then, the probability that an employee's age is in the range $\{22,23,24\}$
can be computed as follow:
\begin{eqnarray*}
 &  & \pmeas\left(\left\{ i\middle|\textrm{Age}(M_{i})\in\{22,23,24\}\right\} \right)\\
 & = & \pmeas\left(\left\{ i=1\right\} \right)\pmeas\left(\left\{ \textrm{Age}(M_{1})\in\{22,23,24\}\right\} \right)\\
 &  & +\pmeas\left(\left\{ i=2\right\} \right)\pmeas\left(\left\{ \textrm{Age}(M_{2})\in\{22,23,24\}\right\} \right)\\
 &  & +\pmeas\left(\left\{ i=3\right\} \right)\pmeas\left(\left\{ \textrm{Age}(M_{3})\in\{22,23,24\}\right\} \right)\\
 & = & \frac{1}{3}\cdot1+\frac{1}{3}\cdot\frac{1}{3}+\frac{1}{3}\cdot0=\frac{4}{9}
\end{eqnarray*}
\item Assume we don't know the probability distributions within the possible
range $D_{1}$, $D_{2}$, and $D_{3}$. All we know is whether they
are \emph{possible} or \emph{necessary}. Then we replace the above
computation from exact value to interval and consider
\begin{eqnarray*}
 &  & \pmeas\left(\left\{ i\middle|\textrm{Age}(M_{i})\in\{22,23,24\}\right\} \right)\\
 & = & \pmeas\left(\left\{ i=1\right\} \right)\pmeas\left(\left\{ \textrm{Age}(M_{1})\in\{22,23,24\}\right\} \right)\\
 &  & +\pmeas\left(\left\{ i=2\right\} \right)\pmeas\left(\left\{ \textrm{Age}(M_{2})\in\{22,23,24\}\right\} \right)\\
 &  & +\pmeas\left(\left\{ i=3\right\} \right)\pmeas\left(\left\{ \textrm{Age}(M_{3})\in\{22,23,24\}\right\} \right)\\
 & = & \frac{1}{3}\cdot[1,1]+\frac{1}{3}\cdot[0,1]+\frac{1}{3}\cdot[0,0]=\left[\frac{1}{3},\frac{2}{3}\right]
\end{eqnarray*}
\yutsung{Check the conditional probability rule in the Dempster-Shafer
Theory!}
\end{itemize}
